---
title: "Project - MPG Data Visualization"
author: "Phatchara Soroschokchai"
date: "2023-12-11"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Data Visualization

## Import Library

```{r, message=FALSE}
library("tidyverse") # for manipulate data (ggplot in this library)
library("ggthemes") # for setting theme
library("patchwork") # multiple plot in one chart
```

## Explore the mpg data

```{r, echo=FALSE}
data("mpg")
mpg$cyl <- as.character(mpg$cyl)
mpg <- mpg %>%
  mutate(trans_label = if_else(grepl("^auto", trans), "automatic", "manual"))
head(mpg)
```

## Data Dictionary

Variable      | Description
------------- | -----------
manufacturer  | manufacturer name
model         | model name
displ         | engine displacement, in litres
year          | year of manufacture
cyl           | number of cylinders
trans         | type of transmission
drv           | the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
cty           | city miles per gallon
hwy           | highway miles per gallon
fl            | fuel type
class         | "type" of car
trans_label   | group of trans

## Top 10 manufacturer with the most average city and highway miles per galon. 
```{r, fig.height=6}
manu_mean_cty <- mpg %>%
  group_by(manufacturer) %>%
  summarise(mean_cty = mean(cty)) %>%
  arrange(desc(mean_cty)) %>%
  head(10)

manu_mean_hwy <- mpg %>%
  group_by(manufacturer) %>%
  summarise(mean_hwy = mean(hwy)) %>%
  arrange(desc(mean_hwy)) %>%
  head(10)

bar1 <- ggplot(manu_mean_cty, 
               mapping=aes(x=mean_cty, y=reorder(manufacturer, -mean_cty))) +
  geom_col(fill = "cadetblue") +
  scale_x_continuous(limits = c(0, 35),
                     oob = scales::squish) +
  labs(title = "Top 10 Manufacturer with the most average city miles per galon",
       x = "Manufacturer",
       y = "Avereage of City Miles per Galon",
       caption = "source: mpg dataset in R") + 
  theme_minimal()

bar2 <- ggplot(manu_mean_hwy, 
               mapping=aes(x=mean_hwy, y=reorder(manufacturer, -mean_hwy))) +
  geom_col(fill = "seagreen4") +
  scale_x_continuous(limits = c(0, 35),
                     oob = scales::squish) +
  labs(title = "Top 10 Manufacturer with the most average highway miles per galon",
       x = "Manufacturer",
       y = "Avereage of Highway Miles per Galon",
       caption = "source: mpg dataset in R") + 
  theme_minimal()

bar1 / bar2
```
Fig. 1. You can see Honda and Volkswagen manufacturer produce the most fuel-efficient cars when measured in miles per gallon on average. If the car of Honda and Volkswagen manufacturer is driven on the highway and in the city, then Honda and Volkswagen are the best options.

## Distribution of city & highway miles per galon by type of drive train

```{r, fig.height=10, fig.width=15}
box1 <- ggplot(mpg, 
               mapping=aes(x=cty, group=drv, col=drv)) +
  geom_boxplot() +
  labs(title = "Box plot of city miles per galon by type of drive train",
       x = "City Miles per Galon",
       y = "Ntype of drive train",
       caption = "source: mpg dataset in R") + 
  theme_minimal()

box2 <- ggplot(mpg, 
               mapping=aes(x=hwy, group=drv, col=drv)) +
  geom_boxplot() +
  labs(title = "Box plot of city miles per galon by type of drive train",
       x = "Highway Miles per Galon",
       y = "type of drive train",
       caption = "source: mpg dataset in R") + 
  theme_minimal()

hist1 <- ggplot(mpg, 
                mapping=aes(x=cty, fill=drv, col=drv)) +
  geom_histogram(bins=20, alpha=0.4, position = "identity") +
  labs(title = "Histogram of highway miles per galon by type of drive train",
       x = "City Miles per Galon",
       y = "Frequency",
       caption = "source: mpg dataset in R") + 
       theme_minimal()

hist2 <- ggplot(mpg, 
                mapping=aes(x=hwy, fill=drv, col=drv)) +
  geom_histogram(bins=20, alpha=0.4, position = "identity") +
  labs(title = "Histogram of highway miles per galon by type of drive train",
       x = "Highway Miles per Galon",
       y = "Frequency",
       caption = "source: mpg dataset in R") + 
       theme_minimal()

(box1 + box2) / (hist1 + hist2)
```
Fig. 2. This chart shows the distribution of miles per galon if a car is driven on a highway or in a city. In conclusion, front-wheel drive cars can get better miles per galon, showing that front-wheel drive cars can save on fuel more than 4-wheel and rear-wheel drive cars.

## Comparison between City and Highway Miles per Galon by number of cylinders.

```{r, fig.width=10, warning=FALSE}
ggplot(mpg, 
       mapping = aes(x=cty, y=hwy, col=cyl)) +
  geom_point(alpha=0.4, size=2) +
  geom_smooth(method="lm", se=FALSE) +
  labs(title = "Comparison between City and Highway Miles per Galon by number of cylinders",
       x = "City Miles per Galon",
       y = "Highway Miles per Galon",
       caption = "source: mpg dataset in R") +
  theme_minimal()
```
Fig. 3. This chart shows a comparison between City MPG and Highway MPG segments based on number of cylinders. It can be seen that the number of cylinders = 4 can make City MPG and Highway MPG mostly very valuable. And if the number of cylinders is = 8, City MPG and Highway MPG can mostly be low.

## Fraction of Car types by Manufaturer

```{r, fig.width=10}
ggplot(mpg, 
       aes(manufacturer, fill=class)) +
  geom_bar(position="fill") + 
  scale_fill_brewer(palette = "PuBuGn") +
  labs(title = "Fraction of Car types by Manufaturer",
       x = "Manufacturer",
       y = "Fraction",
       caption = "source: mpg dataset in R") +
  theme_minimal()
```
Fig. 4. From this graph, you can see that Jeep, Land Rover, Lincoln, and Mercury all produce SUVs cars. Honda only produces mini vans cars. Pontiac only produces mid size cars.

## Box plot of Engine displacement by Transmission and fuel types

```{r, fig.height=10, fig.width=8}
ggplot(mpg, aes(x = displ, group=trans_label, fill=trans_label)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Box plot of Engine displacement by Transmission and fuel types",
       x = "Engine displacement (litres)",
       y = "Transmission (Auto, Manual)") +
  facet_wrap(~ fl, nrow=5)
```
Fig. 5. Automatic transmission cars Always have a larger average cylinder volume (liters) than a car with a manual transmission.



